### 2021-10-12 chapters/introduction.tex
Updates from Overleaf

% ***************************************************************************************************************************************************************
    \section{Gravitational waves} 
    In 2017, the LIGO and Virgo collaborations observed the first direct gravitational waves from a binary neutron star merger, GW170817~\cite{GW170817Detection}.
    This began the era of multi-messenger gravitational-wave observations with coincident detections in many parts of the electromagnetic spectrum~\cite{Goldstein2017,Coulter2017,Troja2017,Nicholl2017,Chornock2017,Margutti2017,Alexander2017,GW170817multi}. \par

    Gravitational waves are a phenomenon predicted by general relativity which was formulated more than 100 years ago by Albert Einstein with the famous formula
    \begin{align}
        G_{\mu\nu} & = R_{\mu\nu} - \frac{1}{2}R  g_{\mu\nu}  =  \frac{8\pi G}{c^4} T_{\mu\nu}, \label{eq:Intro:GeqT}
    \end{align}
    where $G_{\mu\nu}$ is the space-time curvature tensor and $T_{\mu\nu}$ is the stress energy tensor. The local metric tensor is $g_{\mu\nu}$, $R_{\mu\nu}$ is the Ricci tensor and $R$ is the Ricci scalar. 
    The gravitational constant is $G$ and the speed of light in vacuum is $c$, though for remainder of this work we will use geometric units of $G=c=1$ unless otherwise stated. 
    Equation~\ref{eq:Intro:GeqT} demonstrates how the curvature of the space-time, $G_{\mu\nu}$, alters the dynamics of matter, $T_{\mu\nu}$, and vice-versa. \par
    Movement of matter in space can generate gravitational waves.
    For simulation purposes, the movement of matter can be modelled by the Einstein field equations~Eq.(\ref{eq:Intro:GeqT}) and perturbations of black holes~\cite[e.g.,][]{Moncrief1974,Abrahams1996,Andrade1999,Nagar2005}.
    This methodology allows the calculation of multipole functions of odd ($\mathcal{Q}^\times_{lm}$) and even ($\mathcal{Q}^+_{lm}$) parity, with angular indices of $(l,m)$.
    These multipoles can be evaluated by numerical-relativity simulations which allow the extraction of gravitational waves.
    The plus, $h_{+}$, and cross, $h_{\times}$, polarisations of the gravitational waves are related to the multipoles as follows~\cite{Moncrief1974,Abrahams1996,Andrade1999,Nagar2005,Baiotti2008}:\footnote{Please note that $\mathcal{Q}^+_{lm}$ and $\mathcal{Q}^\times_{lm}$ have different units, see Ref.~\cite{Nagar2005} for more information.}
     \begin{align}
        h_{+} - i h_{\times} & = \frac{1}{r\sqrt{2}}\sum_{l,m} {}_{-2}Y^{lm}\left(\mathcal{Q}^{+}_{lm} -i \int_{-\infty}^t\mathcal{Q}^\times_{lm}(t')dt'\right). \label{eq:Intro:MultipoleExpansion}
    \end{align}   
    Here, $r$ is the distance to the gravitational-wave source and ${}_{-2}Y^{lm}$ are the spin-weighted spherical harmonics with spin $s=-2$ and angular indices of $(l,m)$.
    For most sources, gravitational waves are predominantly emitted in the $l=m=2$ mode. 
    In this case Eq.(\ref{eq:Intro:MultipoleExpansion}) evaluates to
     \begin{align}
        h_{+} + i h_{\times} & = \frac{1}{2 r} \sqrt{\frac{5}{2\pi }} e^{2 i \phi } \cos ^4\left(\frac{\theta }{2}\right)\left(\mathcal{Q}^{+}_{22} -i \int_{-\infty}^t\mathcal{Q}^\times_{22}(t')dt'\right), \label{eq:Intro:MultipoleExpansionlm22}
    \end{align}
    where  $\theta$ is the polar angle and $\phi$ is the azimuthal angle usually defined relative to orbital angular momentum vector. \par

@@

    Another binary neutron star (possibly neutron star - black hole) merger, GW190425, was detected in 2019~\cite{GW190425Detection,GWTC2,PopGWTC2}.  However, the luminosity distance of GW190425, $159^{+69}_{-72}\,$Mpc, and the broad sky localisation of the source, $8284\,\mathrm{deg}^2$, precluded any coincident electromagnetic observations~\cite{GW190425Detection}. In comparison, the sky localisation of GW170817 was $28\,\mathrm{deg}^2$ at a distance of $\sim 40\,\Mpc$.
    Therefore, to date, GW170817 is the only gravitational-wave signal that was observed with coincident electromagnetic radiation and will be discussed further in Section~\ref{sec:Intro:NeutronStarMergers}.

@@

% ***************************************************************************************************************************************************************
    \section{Neutron stars}
    \label{sec:Intro:NeutronStars}
    Neutron stars are compact stellar remnants that are formed following supernovae of stars with masses of $\gtrsim 8\,\mathrm{M}_\odot$~\cite[e.g.,][]{Lattimer2004}.
    They can also be formed from the dynamics of binary star systems~\cite[e.g.,][]{Ruiter2019,LiuX2020,LiuD2020}.
    Neutron stars are created when the gravitational collapse of the inner stellar core overcomes the electron degeneracy pressure.
    The proto-neutron star matter is then supported against further collapse by neutron (or quark) degeneracy pressure. \par

@@

% \par

    % Another unique property of neutron stars is that they are opaque to neutrinos.

% ***************************************************************************************************************************************************************
    \section{Neutron star mergers} \label{sec:Intro:NeutronStarMergers}
    Binary neutron star mergers emit vast amounts of energy in gravitational waves.
    The estimated radiated energy for the GW170817 merger was $\gtrsim 5\times 10^{52}\,$ergs~\cite{GW170817Detection}, equivalent to $\sim 1\%$ of the total rest mass of the two neutron stars.
    For a source at a distance of $D$ from the observer, the amplitude of gravitational waves are  $\propto 1/D$, whereas the amplitude of electromagnetic waves are $\propto 1/D^2$.
    Additionally, electromagnetic waves can be attenuated by intervening media between the source and the detector whereas gravitational waves are transparent to the intervening media.
    Hence gravitational waves are detectable over much larger distances in comparison to electromagnetic radiation with the same source luminosity. \par
    Gravitational-wave strain emitted from the inspiral of binary neutron star mergers, approximated by Eqs.(\ref{eq:Intro:FreqEvolution}\Hyphdash*\ref{eq:Intro:chirp_ht}), can be recast in the frequency domain using the stationary phase approximation~\cite{Cutler1994}
    \begin{align}
        h_{22}(f) & \equiv \frac{\mathcal{A}}{D}\mathcal{M}^{5/6}f^{-7/6} \exp(i\Psi(f)), \label{eq:Intro:StationaryPhaseApproxH22} \\
        \Psi(f) & = 2\pi f t_c - \phi_c - \frac{\pi}{4} +\frac{3}{4}\left(8\pi\mathcal{M}f\right)^{-5/3}, \label{eq:Intro:StationaryPhaseApproxPsi} 
    \end{align}
    where $h_{22}(f)$ is the Fourier transform of the $(l,m)=(2,2)$ gravitational-wave strain, and $\Psi(f)$ is the corresponding phase.
    Here, $f$ is the frequency~(Eq.(\ref{eq:Intro:FreqEvolution})), $\phi_c$ and $t_c$ are the phase and time of coalescence, and $\mathcal{A}$ is a factor that depends on the orientation of the system.\footnote{The orientation factor, $\mathcal{A}$, depends on angles intrinsic and extrinsic to the binary system. Intrinsic angles include the angles defining the relative orientation of the neutron star spins to the orbital plane. Extrinsic angles include the angles defining the relative orientation between the orbital plane and each of the detectors.}
    Equations~\ref{eq:Intro:StationaryPhaseApproxH22}\Hyphdash*\ref{eq:Intro:StationaryPhaseApproxPsi} are valid for $a \gtrsim 6M$ where $a$ is the orbital separation of the binary system and $a\sim 6M$ is an approximation of the innermost stable circular orbit~\cite{Cutler1994}.
    Solving Eqs.(\ref{eq:Intro:OrbitalEvolution}\Hyphdash*\ref{eq:Intro:FreqEvolution}) for $a \sim 6M$ gives $f_{max}\approx c^3 ( 6^{3/2}\pi G M )^{-1}\,\mathrm{Hz}$.
    \par
    % The base post-Newtonian expansion parameter is $\sqrt{M/r}$ but it is more convenient to use $\left(\pi M f\right)^{1/3}$ as the post-Newtonian expansion parameter.
    % These two parameters are equal up to and excluding terms of order $(M/r)^{3/2}$.
    The validity for these equations can be extended to times closer to the merger by adding relativistic corrections calculated by post-Newtonian expansion and effective-one-body dynamics~\cite{Cutler1994,Blanchet1995, Buonanno1999,Droz1999,Damour2012,Yagi2013,Blanchet2014}. % to Eq.~\ref{eq:Intro:StationaryPhaseApproxPsi}
    These corrections introduce dependence on the masses, spins, and tidal deformabilities of the individual neutron stars.
    This allows the measurement of these system properties in addition to the chirp mass, given a gravitational-wave inspiral detection.
    However, the accuracy of the individual neutron star properties is significantly less than the accuracy of the chirp mass because the relativistic corrections required to measure these properties are fairly weak. 
    This can be seen by looking at the parameters measured from GW170817.

@@

    Numerical-relativity simulations are required for determining the dynamics of the late inspiral of the binary neutron star and the post-merger remnant.
    To perform a numerical-relativity simulation, the equation of state must be defined, a set of initial conditions must be evaluated in general relativity, and then Einstein's field equations must be evolved.
    Each of these three requirements make these simulations difficult.
    The equation of state for a `normal' neutron star is still unknown, and the equation of state for a hot post-merger remnant is even more uncertain, after all, this is the hottest and densest bulk matter in the Universe.\par

    The initial conditions for binary neutron star mergers in general relativity are complex and a number of methods have been developed for this task.
    Codes that have been used for this include \texttt{LORENE}~\cite{Gourgoulhon2001} and \texttt{COCAL}~\cite{Tsokaros2015}.
    The \texttt{LORENE} formulation assumes a conformally-flat approximation for  neutron star binaries and uses spectral decomposition to solve for the initial conditions~\cite{Gourgoulhon2001}.
    This formulation constructs quasiequilibrium configurations which are approximations of the inspiral dynamics using exactly circular orbits.
    Quasiequilibrium configurations are also generated by \texttt{COCAL}.
    This is performed by excising regions around the neutron stars and solving Green's integral formula using multipole expansion and finite-difference methods~\cite{Uryu2012,Tsokaros2015}.
    Once the initial conditions have been approximated then the space-time can be evolved.

    Space-time evolution methods include the following formulations: Arnowitt-Deser-Misner (ADM)~\cite{Arnowitt2008}\footnote{The original book reference for ADM formalism is out of print and the authors have listed the cited paper as an `intentially unretouched' alternative.},  Baumgarte-Shapiro-Shibata-Nakamura-Oohara-Kojima (BSSNOK)~\cite{Nakamura1987,Shibata1995,Baumgarte1999,Alcubierre2000} and Z4c~\cite{Bona2003,Bona2004,Gundlach2005,Bernuzzi2010,Hilditch2013}. 
    The ADM formalism is a Hamiltonian method that splits the Einstein's field equations into separate time and space components.
    This is achieved by applying a gauge that consists of a shift vector and lapse function~\cite{Arnowitt2008}.
    Unfortunately, ADM numerical simulations were found to be unstable to accumulated errors over the simulation time~\cite{Brandt2000,Kidder2001,Alcubierre2001a} and a number of different formulations were developed to improve this. \par

    To aid computation convergence, the BSSNOK method extended the ADM formalism by introducing a conformal factor to the three-metric and selecting a dissipative hyperbolic driver for the  shift vector~\cite{Nakamura1987,Shibata1995,Nakamura1999,Shibata1999,Baumgarte1999,Alcubierre2000,Alcubierre2001}. % Dissipative hyperbolic driver from Alcubierre2001
    The Z4c formalism modified the BSSNOK method by adding a four-vector of constraints, $Z_\mu$, to the Einstein field equations~\cite{Bona2003,Bona2004,Gundlach2005,Bernuzzi2010,Hilditch2013}. 
    This resulted in a propagating Hamiltonian constraint which helps to reduce constraint-violation growth by damping all constraint violations  except for constant modes~\cite{Gundlach2005}.
    Constraint violations can be more easily tracked as non-vanishing values of $Z_\mu$ though particular attention needs to be paid to reflection of constraint violations at boundaries~\cite{Bernuzzi2010,Ruiz2011,Hilditch2013}.  \par

@@

    Most postmerger numerical-relativity simulations last for only tens of milliseconds, though some longer simulations have been performed between 50-100\,ms in length~\cite{Rezzolla2010,Ciolfi2017,Ciolfi2019,DePietri2020}.
    Although Eq.(\ref{eq:Intro:GWAngularMomentumLoss}) shows an estimated emission time for gravitational waves of $\sim 150\,$ms, the uncertainty in the equation of state could place $f_2$ as low as $\sim 2.2\,\kHz$ and some simulations show that the ellipticity reduces over time~\cite{Takami2015,Shibata2017a}.
    Given these possibilities, a four-fold decrease in $e$ together with a reduction in $f_2$ to $2.2\,\kHz$ would result in $t_{\mathrm{GW}} \sim 12\,$s.
    \redtext{Although the numerical-relativity simulations do not cover enough time period after the merger, they are vital for extracting the gravitational-wave strain in the early post-merger stage where the amplitude of the strain is at its loudest.}

@@

    Two additional oscillation modes were found in simulations that were caused by the interaction between the main f\Hyphdash*mode oscillation ($m=2$) and the $m=0$ radial pressure mode~\cite{Stergioulas2011}.
    This results in a modulated signal that should be present in the frequency response at frequencies $\fpeak \pm f_{m=0}$, designated as $\ftwominuszero$ and $\ftwopluszero$, respectively.
    The $\ftwominuszero$ peak was found to be particularly strong in spectra for high-mass mergers as long as prompt collapse does not occur~\cite{Bauswein2015}.
    Another oscillation mode, labelled $\fspiral$, where $\ftwominuszero < \fspiral < \fpeak$~\cite{Bauswein2015}, was found in the spectra for low-mass mergers.
    This mode was generated by two antipodal bulges at the outer edge of the merger remnant that rotated at a slower frequency than $\fpeak$.
    The antipodal bulges were, in turn, caused by strong tidal deformation in the early post-merger stage.
    For higher-mass mergers that do not experience prompt collapse, the peaks associated with $\ftwominuszero$ dominate over peaks associated with $\fspiral$ in the spectra~\cite{Bauswein2015}.
    In the mass region between these extremes, both $\fspiral$ and $\ftwominuszero$ can be present in the spectra of these post-merger remnants. \par

@@

    A number of challenges are presented by the gravitational-wave strain generated with numerical-relativity simulations.
    Phase errors between successive resolutions accumulate over time and it is not unusual to see phase differences of $\sim 1$\,rad by the time of the merger for the finest resolutions~\cite[e.g.,][]{Radice2014b}.
    Reference~\cite{Takami2015} examined the role of the spatial resolution on the instantaneous phase of the gravitational-wave strain and found a phase difference of $0.9$\,rad between the the finest resolution tests.
    This results in a loss of $\sim 40\%$ of the signal-to-noise compared to a perfectly matched signal when using a matched-filter detection. \par

@@

    We find that we need a network of Einstein Telescope with Cosmic Explorer to detect a post-merger that collapses $\sim10\,\ms$ after the merger.
    Two A+ detectors at design sensitivity can measure a post-merger remnant that collapses $\sim\!10\,\mathrm{ms}$  after the merger to a distance of $\lesssim 10\,\Mpc$.
    If the proposed Neutron star Extreme Matter Observatory is operational, then this distance increases to $\sim\!18\Hyphdash*26\,\Mpc$ thereby increasing the effective volume and hence detection rate by a factor of $\sim 30$.
\par

>>>>

    \section{Gravitational waves} 
    In 2017, the LIGO and Virgo collaborations observed the first direct gravitational waves from a binary neutron star merger, GW170817~\cite{GW170817Detection}.
    This began the era of multi-messenger gravitational-wave observations with coincident detections in many parts of the electromagnetic spectrum~\cite{Goldstein2017,Coulter2017,Troja2017,Nicholl2017,Chornock2017,Margutti2017,Alexander2017,GW170817multi}. \par

    Gravitational waves are a phenomenon predicted by general relativity which was formulated more than 100 years ago by Albert Einstein with the famous formula
    \begin{align}
        G_{\mu\nu} & = R_{\mu\nu} - \frac{1}{2}R  g_{\mu\nu}  =  \frac{8\pi G}{c^4} T_{\mu\nu}, \label{eq:Intro:GeqT}
    \end{align}
    where $G_{\mu\nu}$ is the space-time curvature tensor and $T_{\mu\nu}$ is the stress energy tensor. The local metric tensor is $g_{\mu\nu}$, $R_{\mu\nu}$ is the Ricci tensor and $R$ is the Ricci scalar. 
    The gravitational constant is $G$ and the speed of light in vacuum is $c$, though for remainder of this work we will use geometric units of $G=c=1$ unless otherwise stated. 
    Equation~\ref{eq:Intro:GeqT} demonstrates how the curvature of the space-time, $G_{\mu\nu}$, alters the dynamics of matter, $T_{\mu\nu}$, and vice-versa. \par
    Movement of matter in space can generate gravitational waves.
    For simulation purposes, the movement of matter can be modelled by the Einstein field equations~Eq.(\ref{eq:Intro:GeqT}) and perturbations of black holes~\cite[e.g.,][]{Moncrief1974,Abrahams1996,Andrade1999,Nagar2005}.
    This methodology allows the calculation of multipole functions of odd ($\mathcal{Q}^\times_{lm}$) and even ($\mathcal{Q}^+_{lm}$) parity, with angular indices of $(l,m)$.
    These multipoles can be evaluated by numerical-relativity simulations which allow the extraction of gravitational waves.

    The plus, $h_{+}$, and cross, $h_{\times}$, polarisations of the gravitational waves are related to the multipoles as follows~\cite{Moncrief1974,Abrahams1996,Andrade1999,Nagar2005,Baiotti2008}:\footnote{Please note that $\mathcal{Q}^+_{lm}$ and $\mathcal{Q}^\times_{lm}$ have different units, see Ref.~\cite{Nagar2005} for more information.}
     \begin{align}
        h_{+} - i h_{\times} & = \frac{1}{r\sqrt{2}}\sum_{l,m} {}_{-2}Y^{lm}\left(\mathcal{Q}^{+}_{lm} -i \int_{-\infty}^t\mathcal{Q}^\times_{lm}(t')dt'\right). \label{eq:Intro:MultipoleExpansion}
    \end{align}   
    Here, $r$ is the distance to the gravitational-wave source and ${}_{-2}Y^{lm}$ are the spin-weighted spherical harmonics with spin $s=-2$ and angular indices of $(l,m)$.
    For most sources, gravitational waves are predominantly emitted in the $l=m=2$ mode. 
    In this case Eq.(\ref{eq:Intro:MultipoleExpansion}) evaluates to
     \begin{align}
        h_{+} + i h_{\times} & = \frac{1}{2 r} \sqrt{\frac{5}{2\pi }} e^{2 i \phi } \cos ^4\left(\frac{\theta }{2}\right)\left(\mathcal{Q}^{+}_{22} -i \int_{-\infty}^t\mathcal{Q}^\times_{22}(t')dt'\right), \label{eq:Intro:MultipoleExpansionlm22}
    \end{align}
    where  $\theta$ is the polar angle and $\phi$ is the azimuthal angle usually defined relative to orbital angular momentum vector. \par

@@

    Another binary neutron star (possibly neutron star - black hole) merger, GW190425, was detected in 2019~\cite{GW190425Detection,GWTC2,PopGWTC2}.

    However, the luminosity distance of GW190425, $159^{+69}_{-72}\,$Mpc, and the broad sky localisation of the source, $8284\,\mathrm{deg}^2$, precluded any coincident electromagnetic observations~\cite{GW190425Detection}. In comparison, the sky localisation of GW170817 was $28\,\mathrm{deg}^2$ at a distance of $\sim 40\,\Mpc$.
    Therefore, to date, GW170817 is the only gravitational-wave signal that was observed with coincident electromagnetic radiation and will be discussed further in Section~\ref{sec:Intro:NeutronStarMergers}.

@@

    \section{Neutron stars}
    \label{sec:Intro:NeutronStars}
    Neutron stars are compact stellar remnants that are formed following supernovae of stars with masses of $\gtrsim 8\,\mathrm{M}_\odot$~\cite[e.g.,][]{Lattimer2004}.
    They can also be formed from the dynamics of binary star systems~\cite[e.g.,][]{Ruiter2019,LiuX2020,LiuD2020}.
    Neutron stars are created when the gravitational collapse of the inner stellar core overcomes the electron degeneracy pressure.
    The proto-neutron star matter is then supported against further collapse by neutron (or quark) degeneracy pressure. \par

@@

    \section{Neutron star mergers} \label{sec:Intro:NeutronStarMergers}
    Binary neutron star mergers emit vast amounts of energy in gravitational waves.
    The estimated radiated energy for the GW170817 merger was $\gtrsim 5\times 10^{52}\,$ergs~\cite{GW170817Detection}, equivalent to $\sim 1\%$ of the total rest mass of the two neutron stars.
    For a source at a distance of $D$ from the observer, the amplitude of gravitational waves are  $\propto 1/D$, whereas the amplitude of electromagnetic waves are $\propto 1/D^2$.
    Additionally, electromagnetic waves can be attenuated by intervening media between the source and the detector whereas gravitational waves are transparent to the intervening media.
    Hence gravitational waves are detectable over much larger distances in comparison to electromagnetic radiation with the same source luminosity. \par

    Gravitational-wave strain emitted from the inspiral of binary neutron star mergers, approximated by Eqs.(\ref{eq:Intro:FreqEvolution}\Hyphdash*\ref{eq:Intro:chirp_ht}), can be recast in the frequency domain using the stationary phase approximation~\cite{Cutler1994}
    \begin{align}
        h_{22}(f) & \equiv \frac{\mathcal{A}}{D}\mathcal{M}^{5/6}f^{-7/6} \exp(i\Psi(f)), \label{eq:Intro:StationaryPhaseApproxH22} \\
        \Psi(f) & = 2\pi f t_c - \phi_c - \frac{\pi}{4} +\frac{3}{4}\left(8\pi\mathcal{M}f\right)^{-5/3}, \label{eq:Intro:StationaryPhaseApproxPsi} 
    \end{align}
    where $h_{22}(f)$ is the Fourier transform of the $(l,m)=(2,2)$ gravitational-wave strain, and $\Psi(f)$ is the corresponding phase.
    Here, $f$ is the frequency~(Eq.(\ref{eq:Intro:FreqEvolution})), $\phi_c$ and $t_c$ are the phase and time of coalescence, and $\mathcal{A}$ is a factor that depends on the orientation of the system.\footnote{The orientation factor, $\mathcal{A}$, depends on angles intrinsic and extrinsic to the binary system. Intrinsic angles include the angles defining the relative orientation of the neutron star spins to the orbital plane. Extrinsic angles include the angles defining the relative orientation between the orbital plane and each of the detectors.}
    Equations~\ref{eq:Intro:StationaryPhaseApproxH22}\Hyphdash*\ref{eq:Intro:StationaryPhaseApproxPsi} are valid for $a \gtrsim 6M$ where $a$ is the orbital separation of the binary system and $a\sim 6M$ is an approximation of the innermost stable circular orbit~\cite{Cutler1994}.
    Solving Eqs.(\ref{eq:Intro:OrbitalEvolution}\Hyphdash*\ref{eq:Intro:FreqEvolution}) for $a \sim 6M$ gives $f_{max}\approx c^3 ( 6^{3/2}\pi G M )^{-1}\,\mathrm{Hz}$.
    \par

    The validity for these equations can be extended to times closer to the merger by adding relativistic corrections calculated by post-Newtonian expansion and effective-one-body dynamics~\cite{Cutler1994,Blanchet1995, Buonanno1999,Droz1999,Damour2012,Yagi2013,Blanchet2014}. 
    These corrections introduce dependence on the masses, spins, and tidal deformabilities of the individual neutron stars.
    This allows the measurement of these system properties in addition to the chirp mass, given a gravitational-wave inspiral detection.
    However, the accuracy of the individual neutron star properties is significantly less than the accuracy of the chirp mass because the relativistic corrections required to measure these properties are fairly weak. 
    This can be seen by looking at the parameters measured from GW170817.

@@

    Numerical-relativity simulations are required for determining the dynamics of the late inspiral of the binary neutron star and the post-merger remnant.
    To perform a numerical-relativity simulation, the equation of state must be defined, a set of initial conditions must be evaluated in general relativity, and then Einstein's field equations must be evolved.
    Each of these three requirements make these simulations difficult.
    The equation of state for a `normal' neutron star is still unknown, and the equation of state for a hot post-merger remnant is even more uncertain, after all, this is the hottest and densest bulk matter in the Universe.

    \par

    The initial conditions for binary neutron star mergers in general relativity are complex and a number of methods have been developed for this task.
    Codes that have been used for this include \texttt{LORENE}~\cite{Gourgoulhon2001} and \texttt{COCAL}~\cite{Tsokaros2015}.
    The \texttt{LORENE} formulation assumes a conformally-flat approximation for  neutron star binaries and uses spectral decomposition to solve for the initial conditions~\cite{Gourgoulhon2001}.
    This formulation constructs quasiequilibrium configurations which are approximations of the inspiral dynamics using exactly circular orbits.
    Quasiequilibrium configurations are also generated by \texttt{COCAL}.
    This is performed by excising regions around the neutron stars and solving Green's integral formula using multipole expansion and finite-difference methods~\cite{Uryu2012,Tsokaros2015}.

    Once the initial conditions have been approximated then the space-time can be evolved.

    Space-time evolution methods include the following formulations: Arnowitt-Deser-Misner (ADM)~\cite{Arnowitt2008}\footnote{The original book reference for ADM formalism is out of print and the authors have listed the cited paper as an `intentially unretouched' alternative.},  Baumgarte-Shapiro-Shibata-Nakamura-Oohara-Kojima (BSSNOK)~\cite{Nakamura1987,Shibata1995,Baumgarte1999,Alcubierre2000} and Z4c~\cite{Bona2003,Bona2004,Gundlach2005,Bernuzzi2010,Hilditch2013}.

    The ADM formalism is a Hamiltonian method that splits the Einstein's field equations into separate time and space components.
    This is achieved by applying a gauge that consists of a shift vector and lapse function~\cite{Arnowitt2008}.
    Unfortunately, ADM numerical simulations were found to be unstable to accumulated errors over the simulation time~\cite{Brandt2000,Kidder2001,Alcubierre2001a} and a number of different formulations were developed to improve this. \par

    To aid computation convergence, the BSSNOK method extended the ADM formalism by introducing a conformal factor to the three-metric and selecting a dissipative hyperbolic driver for the  shift vector~\cite{Nakamura1987,Shibata1995,Nakamura1999,Shibata1999,Baumgarte1999,Alcubierre2000,Alcubierre2001}. 
    The Z4c formalism modified the BSSNOK method by adding a four-vector of constraints, $Z_\mu$, to the Einstein field equations~\cite{Bona2003,Bona2004,Gundlach2005,Bernuzzi2010,Hilditch2013}. 
    This resulted in a propagating Hamiltonian constraint which helps to reduce constraint-violation growth by damping all constraint violations  except for constant modes~\cite{Gundlach2005}.
    Constraint violations can be more easily tracked as non-vanishing values of $Z_\mu$ though particular attention needs to be paid to reflection of constraint violations at boundaries~\cite{Bernuzzi2010,Ruiz2011,Hilditch2013}.  \par

@@

    Most postmerger numerical-relativity simulations last for only tens of milliseconds, though some longer simulations have been performed between 50-100\,ms in length~\cite{Rezzolla2010,Ciolfi2017,Ciolfi2019,DePietri2020}.
    % Rezzolla2010 long early simulation showing conservation of M_adm and J_gw
    Although Eq.(\ref{eq:Intro:GWAngularMomentumLoss}) shows an estimated emission time for gravitational waves of $\sim 150\,$ms, the uncertainty in the equation of state could place $f_2$ as low as $\sim 2.2\,\kHz$ and some simulations show that the ellipticity reduces over time~\cite{Takami2015,Shibata2017a}.
    Given these possibilities, a four-fold decrease in $e$ together with a reduction in $f_2$ to $2.2\,\kHz$ would result in $t_{\mathrm{GW}} \sim 12\,$s.

    Although the numerical-relativity simulations do not cover enough time period after the merger, they are vital for extracting the gravitational-wave strain in the early post-merger stage where the amplitude of the strain is at its loudest.

@@

    Two additional oscillation modes were found in simulations that were caused by the interaction between the main f\Hyphdash*mode oscillation ($m=2$) and the $m=0$ radial pressure mode~\cite{Stergioulas2011}.

    This results in a modulated signal that should be present in the frequency response at frequencies $\fpeak \pm f_{m=0}$, designated as $\ftwominuszero$ and $\ftwopluszero$, respectively.
    %\par
    The $\ftwominuszero$ peak was found to be particularly strong in spectra for high-mass mergers as long as prompt collapse does not occur~\cite{Bauswein2015}.
    Another oscillation mode, labelled $\fspiral$, where $\ftwominuszero < \fspiral < \fpeak$~\cite{Bauswein2015}, was found in the spectra for low-mass mergers.
    This mode was generated by two antipodal bulges at the outer edge of the merger remnant that rotated at a slower frequency than $\fpeak$.
    The antipodal bulges were, in turn, caused by strong tidal deformation in the early post-merger stage.
    For higher-mass mergers that do not experience prompt collapse, the peaks associated with $\ftwominuszero$ dominate over peaks associated with $\fspiral$ in the spectra~\cite{Bauswein2015}.
    In the mass region between these extremes, both $\fspiral$ and $\ftwominuszero$ can be present in the spectra of these post-merger remnants. \par

@@

    A number of challenges are presented by the gravitational-wave strain generated with numerical-relativity simulations.
    Phase errors between successive resolutions accumulate over time and it is not unusual to see phase differences of $\sim 1$\,rad by the time of the merger for the finest resolutions~\cite[e.g.,][]{Radice2014b}.

    Reference~\cite{Takami2015} examined the role of the spatial resolution on the instantaneous phase of the gravitational-wave strain and found a phase difference of $0.9$\,rad between the the finest resolution tests.
    This results in a loss of $\sim 40\%$ of the signal-to-noise compared to a perfectly matched signal when using a matched-filter detection. \par

@@

    We find that we need a network of Einstein Telescope with Cosmic Explorer to detect a post-merger that collapses $\sim10\,\ms$ after the merger.
    Two A+ detectors at design sensitivity can measure a post-merger remnant that collapses $\sim\!10\,\mathrm{ms}$  after the merger to a distance of $\lesssim 10\,\Mpc$.
    If the proposed Neutron star Extreme Matter Observatory is operational, then this distance increases to $\sim\!18\Hyphdash*26\,\Mpc$ thereby increasing the effective volume and hence detection rate by a factor of $\sim 30$.
